A comparative analysis of AquaMaps and IUCN species distribution maps

### 2015-12-11 R/data_explore_pt2.Rmd
script uses new data
I think the distribution by taxonomic group code is working now...

spp_list <- spp_list %>%
  mutate(lg_area = ifelse(am_area > iucn_area, am_area, iucn_area),
         sm_area = ifelse(am_area < iucn_area, am_area, iucn_area),
         lg_area_pct = 100 * lg_area / max(total_area, na.rm = TRUE)) ### use largest data set as approx for total ocean range

@@

>>>>

@@

###(4) Looking at the percent of smaller species range that is within the larger range.

For each range

```{r pct_small_within_large, echo = FALSE, warning = FALSE, eval = TRUE}

spp_df4 <- spp_list %>%
  dplyr::select(sciname, sm_perc, sm_range) %>%
  unique()

hist_sm_within_lg <- ggplot(spp_df4, aes(sm_perc, fill = sm_range)) + 
  ggtheme_plot + 
  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 4, alpha = .8) +
  scale_fill_discrete(name = 'Dataset with smaller range', labels = c('AquaMaps', 'IUCN')) + 
  labs(title = 'Percent of smaller range within larger range',
       x = 'Percent of smaller range that falls within the larger range', 
       y = sprintf('Proportion of all species (n = %s)', nrow(spp_df4)))

ggsave(file.path(dir_fig, 'PercSmallRangeInLarger.png'), width = 6, height = 4)

print(hist_sm_within_lg)

### 2015-12-11 data_explore.Rmd
I think the distribution by taxonomic group code is working now...

###(4) Looking at the percent of smaller species range that is within the larger range.

For each range

```{r pct_small_within_large, echo = FALSE, warning = FALSE, eval = TRUE}

spp_df4 <- spp_list %>%
  dplyr::select(sciname, sm_perc, sm_range) %>%
  unique()

hist_sm_within_lg <- ggplot(spp_df4, aes(sm_perc, fill = sm_range)) + 
  ggtheme_plot + 
  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 4, alpha = .8) +
  scale_fill_discrete(name = 'Dataset with smaller range', labels = c('AquaMaps', 'IUCN')) + 
  labs(title = 'Percent of smaller range within larger range',
       x = 'Percent of smaller range that falls within the larger range', 
       y = sprintf('Proportion of all species (n = %s)', nrow(spp_df4)))

ggsave(file.path(dir_fig, 'PercSmallRangeInLarger.png'), width = 6, height = 4)

print(hist_sm_within_lg)

@@

Plotting species from q1 (well-matched) (see data_explore_quadmaps.html for plots)
```{r plot_spp_from_q1, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}

x <- 20
y <- nrow(spp_list_q1)

spp_names_q1 <- spp_list_q1$sciname[seq(1, y, y/x)]

for (spp in spp_names_q1) { # i = 1
  spp_plot <- plot_rangemap(spp)
  spp_solo <- spp_list %>% filter(sciname == spp)
  text_tag <- sprintf('Overlap: %s%%, area ratio: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
  spp_plot <- spp_plot +
    labs(title = sprintf('%s, quadrant 1: %s', spp, text_tag))

  print(spp_plot)

  # ggsave(filename = file.path(dir_fig, sprintf('rangemap_q1_%s.png', str_replace(tolower(spp), ' ', '_'))),
  #        plot = spp_plot, 
  #        width = 8, height = 4, dpi = 200)
}

Plotting species from q2 (distribution-matched)
```{r plot_spp_from_q2, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}

x <- 20
y <- nrow(spp_list_q2)

spp_names_q2 <- spp_list_q2$sciname[seq(1, y, y/x)]

for (spp in spp_names_q2) { # i = 1
  spp_plot <- plot_rangemap(spp)
  spp_solo <- spp_list %>% filter(sciname == spp)
  text_tag <- sprintf('Overlap: %s%%, area ratio: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
  spp_plot <- spp_plot +
    labs(title = sprintf('%s, quadrant 2: %s', spp, text_tag))
  print(spp_plot)

  # ggsave(filename = file.path(dir_fig, sprintf('rangemap_q2_%s.png', str_replace(tolower(spp), ' ', '_'))),
  #        plot = spp_plot, 
  #        width = 8, height = 4, dpi = 200)
}

Plotting species from q3 (area-matched)
```{r plot_spp_from_q3, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}

x <- 20
y <- nrow(spp_list_q3)

spp_names_q3 <- spp_list_q3$sciname[seq(1, y, y/x)]

for (spp in spp_names_q3) { # i = 1
  spp_plot <- plot_rangemap(spp)
  spp_solo <- spp_list %>% filter(sciname == spp)
  text_tag <- sprintf('Overlap: %s%%, area ratio: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
  spp_plot <- spp_plot +
    labs(title = sprintf('%s, quadrant 3: %s', spp, text_tag))
  print(spp_plot)

  # ggsave(filename = file.path(dir_fig, sprintf('rangemap_q3_%s.png', str_replace(tolower(spp), ' ', '_'))),
  #        plot = spp_plot, 
  #        width = 8, height = 4, dpi = 200)
}

Plotting species from q4 (poorly matched)
```{r plot_spp_from_q4, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}

x <- 20
y <- nrow(spp_list_q4)

spp_names_q4 <- spp_list_q4$sciname[seq(1, y, y/x)]

for (spp in spp_names_q4) { # i = 1
  spp_plot <- plot_rangemap(spp)
  spp_solo <- spp_list %>% filter(sciname == spp)
  text_tag <- sprintf('Overlap: %s%%, area ratio: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
  spp_plot <- spp_plot +
    labs(title = sprintf('%s, quadrant 4: %s', spp, text_tag))
  print(spp_plot)

  # ggsave(filename = file.path(dir_fig, sprintf('rangemap_q4_%s.png', str_replace(tolower(spp), ' ', '_'))),
  #        plot = spp_plot, 
  #        width = 8, height = 4, dpi = 200)
}

>>>>

@@

``` {r child = 'R/data_explore_pt3.Rmd', eval = FALSE}
# Multiple species maps per quadrant
```

### 2015-12-11 ms_outline.Rmd
I think the distribution by taxonomic group code is working now...

---
title: "A comparative analysis of AquaMaps and IUCN species ranges"
author: "JA, CO, BH"
output:
  html_document:
    toc: true
    number_sections: true
    theme: "spacelab"
    highlight: "haddock" 
---

@@

iucn_maps_file <- 'data/spp_iucn_maps.csv'
spp_iucn_maps <- read.csv(iucn_maps_file, stringsAsFactors = FALSE)

@@

@@

The maps provided by the IUCN for each assessed species are created by one or more experts on that species. There are general guidelines to how the map should be created. Specifically, the final mapped polygons are meant to indicate that a species *probably only occurs within the polygon* but are *not* meant to indicate that the species is evenly distributed everywhere within the polygon. The guidelines vary slightly for each type of species (terrestrial, freshwater, marine). These maps are typically dependent on observation records from databases such as GBIF.

IUCN only publishes spatial distribution map sets for taxonomic groups in which at least 90% of the species within the taxon group have been evaluated.  While this helps prevent sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.  As of December 2015, IUCN species distribution maps were available for XXX marine species in YYY taxonomic groups.  However, only YYY of these species have been formally assessed under the IUCN Red List of Threatened Species (listed on that spreadsheet, and habitat = "marine" on IUCN species information pages).

<!--
There have been multiple efforts to map marine species based on both observational data and known life history and habitat preferences. AquaMaps and IUCN are two of these efforts.
-->

@@

The IUCN publishes species distribution maps for as collections of shapefiles for a range of taxonomic groups.  Species spatial distributions are determined by experts based upon ___known occurrence, suitable habitat, and blah___.  The IUCN presents these maps as 'limits of distibution' with the caveat that the species 'probably occurs within this polygon, but it does not mean that it is distributed equally within that polygon or occurs everywhere within that polygon'.<cite?>  IUCN only publishes spatial distribution map sets for taxonomic groups in which at least 90% of the species within the taxon group have been evaluated.  While this helps prevent sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.  As of December 2015, IUCN had published species distribution maps for nearly 12,000 marine species across 26 taxonomic groups, though fewer than 4,500 of these have been formally assessed under the IUCN Red List of Threatened Species.

* _something about format - polygons, fields, etc? presence/no presence_
* _presence field, and seasonality info_
* _terrestrial and marine both_
* _use as a reference:_ https://www.conservationtraining.org/course/view.php?id=217&lang=en

<!--
Spatial designation of the species distribution is left to the experts with strict guidance for how to assign the designation. Using known occurrence data, experts depict the species distribution by polygons using a systematic methodology laid out by the [IUCN](http://www.amphibians.org/wp-content/uploads/2013/09/Red-List-Mapping-standards-ARLA-Jan2014-web-version.pdf) (this needs a better source).

__Data Type__  
Shapefiles for individual species. Spatial files include subpopulation information, presence (extinct/extant etc.) and seasonality information (breeding/migration/origin etc.).  
* ___Include other columns typically within each shapefile___

__Methods__  
From [IUCN site](http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources):  
    * The species distribution maps, sometimes referred to as 'limits of distribution' or 'field guide' maps, aim to provide the current known distribution of the species within its native range. The limits of distribution can be determined by using known occurrences of the species, along with the knowledge of habitat preferences, remaining suitable habitat, elevation limited, and other expert knowledge of the species and its range. In most cases the range is depicted as polygons. _A polygon displaying the limits of a species distribution is essentially meant to communicate that the species probably occurs within this polygon, but it does not mean that it is distributed equally within that polygon or occurs everywhere within that polygon_.

* ___Spatial files for IUCN species groups attempt to include a broad sample across the group, to avoid sampling bias (maybe?)___ - but only some species groups have had spatial data released, so some taxa are not represented at all 
    * from http://www.iucnredlist.org/technical-documents/spatial-data: _"This spatial data provided below is for comprehensively assessed taxonomic groups."  "... the data is divided by taxonomic groups.  Groups that have been comprehensively assessed are those containing >90% species evaluated within the described taxon group."_
    -->

@@

[Site](http://www.aquamaps.org/)

AquaMaps (Kaschner et al. 2013) offers species range maps based on modeled relative environmental suitability.  For each assessed species, data from observations and expert knowledge are used to generate bioclimatic envelopes for sea surface temperature, salinity, depth, productivity, and where applicable, sea ice cover and/or distance to land.  The species' suite of bioclimatic envelopes is compared to a map of environmental attributes on a 0.5 degree global grid, creating a global map of ___probability of occurrence___ (or ___relative environmental suitability?___ (see Kaschner et al 2006 and 2011)), with cell values ranging from 0 (unsuitable) to 1.00 (highly suitable).  These maps can then be tweaked by experts altering the parameters and regenerating the modeled map. As of December 2015, nearly 23,000 species have been mapped, though only 5,600 of these have been evaluated for the IUCN Red List. ___Expert mode to further refine models?___ See (Kaschner et al. 2013) for more detailed methods.

@@

<!--
### Spatial resolution - not sure we need this, it is not really _resolution_ but more about nuances of distribution... difference already explained above?

These two differ in spatial resolution - AquaMaps provides higher resolution information about where species do and do not occur globally whereas the IUCN range maps are unable to provide specific information about where within a polygon a species occurs.
-->

#### Poor overlap between assessed species

| Map set   | n~total~                | n~Red_List~                                                               |
| :-------: | ----------------------: | ------------------------------------------------------------------------: |
| AquaMaps  | `r nrow(spp_am)`        | `r nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))`         |
| IUCN      | `r nrow(spp_iucn_maps)` | `r nrow(spp_iucn_maps %>% filter(binomial %in% spp_iucn_marine$sciname))` |
| Both      | `r nrow(spp_am %>% filter(sciname %in% spp_iucn_maps$binomial))` | `r nrow(spp_map_pairs)`          |

* this highlights that each provides valuable and unduplicated spatial range information.

#### Summarize this section?

@@

# In what ways do the two data sets differ

<!--
![](./figures/rangemap_chromis_alpha.png)

__Figure 1:__ Species range map for _Chromis alpha_ from the AquaMaps (orange) and IUCN (purple) datasets.
___is this the right map to use here? find a good one with cool differences___
-->

### Defining and comparing spatial alignment between the two data sets

<!--
___Main messages:___

* Definition of "alignment" in two dimensions, incl "good" vs "poor"
* basic methodology (raster, overlay, calculate)? or should there be a separate methodology section?
-->

<!--
* Note that alignment does not necessarily mean accuracy or precision - a poorly-understood species could appear to be well-aligned.  
-->

@@

<!--
### Determining "presence" for each data set

To compare these two datasets for all species found in each (n = ___1920___?), we rasterized the IUCN polygons to the same cell resolution as AquaMaps: 0.5 degrees in both latitude and longitude. For every grid cell where the polygon overlapped ___the center point? or is it based on ANY overlap, e.g. using weights and including any non-zero weight?  This could boost IUCN presence___, a value of 1 was assigned. For the purposes of comparison, the probability of occurrence values in the AquaMaps data were ignored and all cells where there was a value greater than 0 were assigned a 1. This way we have comparable maps for total species range. 
- ___How do we justify a zero-threshold here?___ 
    - need to understand the "probability" - is it really ___relative environmental suitability___   From Kaschner et al 2011: "Patterns based on a presence threshold RES.0.6 (Fig. 1) were most strongly supported by empirical species richness data (see ‘Validation of species richness’, below)."
    - since IUCN "presence" seems to be pretty loose, set a wide tolerance on AM?
-->

___Main messages:___

* breaking down alignment differences by quadrant provides useful information.
    * q1 = well-matched (well understood? not necessarily, could be luck or over-generous environmental envelopes)
    * other quadrants have defining characteristics that can highlight areas where we should look for better data and better understanding

<!--
Figure 1 shows four representative species: Kajikia Albida, with a high degree of alignment in both area and distribution; Galaxea cryptoramosa, well-matched in spatial alignment but poorly matched in areal alignment; Conus magnificus, well-matched in area but poorly matched in distribution; and Acanthurus nigrosis, poorly matched in both area and distribution.

![](figures/four_maps.png)

___Figure 1: Representative species with distribution mapped in both AquaMaps and IUCN data sets.___  
-->

@@

___Figure 1: Distribution alignment vs area alignment for `r nrow(spp_map_pairs)` species mapped in both Aquamaps and IUCN species distribution maps.___

@@

<!--
(Looking at some representative species maps)
The range described by AquaMaps map covers XXX km^2^; the range described by IUCN map covers YYY km^2^; the overlap between the two maps covers ZZZ km^2^.  Distribution match = ZZZ km^2^ / XXX km^2^; area match = XXX km^2^ / YYY km^2^.

* Single species map: ___Do we want this as a separate figure?___ _it helps to explain what the hell we're talking about; but we could also just jump to the quadrant map?_ 
* ___OR: plot four representative maps FIRST, to show differences in the two dimensions of alignment, THEN show the quadrant map with those species indicated.___ 
    * _This is backwards from how we did the analysis, but I think might flow better for the reader._
-->

@@

___move to discussion___ We were not surprised to find that map pairs for marine mammals, tunas, and billfishes were well-matched (quadrants 1 and 2); many of these species are global-ranging pelagic organisms, with some ranges encompassing nearly the entire ocean.  Mangroves are also not a surprise, being immobile and easily mapped by remote sensing methods.

@@

___Other hypotheses/case studies to explain mechanistic differences?___

@@

### Differences in global distribution between data sets

### Effect of changing threshold on area representation

![](figures/AMThreshold.png)

___Where does the 40% threshold come from?  Is it commonly used?  Justification?___ Show effect of using a 40% threshold vs a 0% threshold.

![](figures/scatter_AMThreshold.png)
Effect of selecting different thresholds on total "present" area - linearish, no big steps or drops to indicate sudden change in tradeoff.  Do regression - slope indicates that sensitivity is less than 1:1

# Methodology? discussion?

>>>>

---
title: "A comparative analysis of AquaMaps and IUCN species distribution maps"
author: "JA, CO, BH"
output:
  html_document:
    toc: true
    number_sections: true
    theme: "spacelab"
    highlight: "haddock" 
---

@@

@@

iucn_maps_file <- 'data/spp_iucn_maps.csv'
spp_iucn_maps <- read.csv(iucn_maps_file, stringsAsFactors = FALSE) %>%
  filter(spp_group != 'hagfishes') %>%           ### not from IUCN...
  filter(sciname %in% spp_iucn_marine$sciname)   ### ditch freshwater spp

@@

Three comparisons presented here:

* Differences in overall species representation - by taxa and by spatial distribution - how well does each set of maps represent the global distribution of species, the global distribution of biodiversity
* alignment between maps for species in common - how well do the maps align, when compared at a species-by-species level
* sensitivity analysis to AquaMaps presence threshold - diff reports use diff thresholds; what is the likely effect? is there a "correct" threshold?
    * this seems like a sub-analysis to justify choice of 0% threshold etc.  
    * ___That graph should be box plots at each threshold, not simply mean and/or median.___

@@

The IUCN publishes species distribution maps for as collections of shapefiles for a range of taxonomic groups, including both terrestrial and marine organisms.  Species spatial distributions are determined by experts based upon known occurrences of the species and expert understanding of range and habitat preferences.  These maps are typically dependent on observation records from databases ___such as GBIF other examples?___, and IUCN provides general guidelines to facilitate consistency in how the maps are generated.

The IUCN presents these maps as 'limits of distibution'; they are meant to indicate that a species probably occurs within the defined polygon, but are not meant to indicate that the species is evenly distributed everywhere within the polygon.  While the polygons roughly define regions of presence/absence, additional attributes provide information on extant/extinct ranges, native/introduced ranges, and seasonality.

IUCN publishes spatial distribution map sets only for taxonomic groups in which at least 90% of the species within the taxon group have been evaluated.  While this mitigates sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.  As of December 2015, IUCN had published species distribution maps for nearly 12,000 marine species across 26 taxonomic groups, though fewer than 4,500 of these have been formally assessed under the IUCN Red List of Threatened Species.

* _use as a reference:_ 
    * https://www.conservationtraining.org/course/view.php?id=217&lang=en
    * http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources

@@

AquaMaps (Kaschner et al. 2013) offers species range maps based on modeled relative environmental suitability.  For each assessed species, data from observations and expert knowledge are used to generate bioclimatic envelopes for sea surface temperature, salinity, depth, productivity, and where applicable, sea ice cover and/or distance to land.  The species' suite of bioclimatic envelopes is compared to a map of environmental attributes on a 0.5 degree global grid, creating a global map of ___probability of occurrence___ (or ___relative environmental suitability?___ (see Kaschner et al 2006 and 2011)), with cell values ranging from 0 (unsuitable) to 1.00 (highly suitable).  These maps can then be tweaked by experts altering the parameters and regenerating the modeled map. As of December 2015, nearly 23,000 species have been mapped, though only 5,600 of these species have been evaluated for the IUCN Red List. ___Expert mode to further refine models?___ See (Kaschner et al. 2013) for more detailed methods.

@@

### Overlap between assessed species

| Map set   | n~total~                | n~Red_List~                                                               |
| :-------: | ----------------------: | ------------------------------------------------------------------------: |
| AquaMaps  | `r nrow(spp_am)`        | `r nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))`         |
| IUCN      | `r nrow(spp_iucn_maps)` | `r nrow(spp_iucn_maps %>% filter(sciname %in% spp_iucn_marine$sciname))`  |
| Both      | `r nrow(spp_am %>% filter(sciname %in% spp_iucn_maps$sciname))` | `r nrow(spp_map_pairs)`           |

@@

# Analysis of taxonomic and regional inclusion

Focus on IUCN red-list evaluated species (incl or not incl DD)? or all species?

## Taxonomic distribution between data sets

Examine the overlapping and non-overlapping species by taxonomic group (order? family?) - are there differences in taxonomic group coverage between the two data sets?

IUCN - translate the taxonomic group into families - is there a way to automate this? http://www.itis.gov/downloads/?
AM - access taxa group information directly from AM species list - order? family?

``` {r breakdown by taxa, echo = FALSE}

sciname_lookup <- read.csv('sciname_lookups/scinames.csv', stringsAsFactors = FALSE)

iucn_maps_class <- spp_iucn_maps %>% 
  ### add sciname list, then jump through hoops to delete dupes
  ### NOTE: this also ditches ones with no sciname match...
  separate(sciname, 'genus', sep = ' ', remove = FALSE, extra = 'drop') %>%
  mutate(genus = tolower(genus)) %>%
  left_join(sciname_lookup, by = 'genus') %>% 
  filter(!(str_detect(tolower(spp_group), 'grass')    & !str_detect(kingdom, 'plant'))) %>%
  filter(!(str_detect(tolower(spp_group), 'mangrove') & !str_detect(kingdom, 'plant'))) %>%
  filter(!(str_detect(tolower(spp_group), 'coral')    & !str_detect(kingdom, 'animal'))) %>%
  filter(!(str_detect(tolower(spp_group), 'reptile')  & !str_detect(kingdom, 'animal'))) %>%
  filter(!(str_detect(tolower(spp_group), 'seabream') & !str_detect(kingdom, 'animal'))) %>%
  filter(!(str_detect(tolower(spp_group), 'wrass')    & !str_detect(kingdom, 'animal'))) %>%
  filter(!(str_detect(tolower(spp_group), 'mamm')     & !str_detect(kingdom, 'animal'))) %>%
  filter(!is.na(phylum)) %>%
  select(-kingdom, -phylum, -family) %>%
  unique()

taxa_iucn_all <- iucn_maps_class[!duplicated(iucn_maps_class$sciname), ] %>%
  filter(!is.na(class)) %>% ### don't count the unknowns
  mutate(class = str_to_title(class),
         n_iucn_tot = n()) %>%        ### length of whole maps list
  group_by(class) %>%
  summarize(n_iucn_class_all  = n(),
            p_iucn_class_all  = round((n_iucn_class_all/first(n_iucn_tot) * 100), 2))

taxa_iucn_solo <- iucn_maps_class[!duplicated(iucn_maps_class$sciname), ] %>%
  filter(!is.na(class)) %>% ### don't count the unknowns
  filter(!sciname %in% spp_am$sciname) %>%  ### filter list to IUCN only
  mutate(class = str_to_title(class),
         n_iucn_solo = n()) %>%       ### length of IUCN only list
  group_by(class) %>%
  summarize(n_iucn_class_solo = n(),
            p_iucn_class_solo = round((n_iucn_class_solo/first(n_iucn_solo) * 100), 2))

taxa_am_all <- spp_am %>%
  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
  mutate(class = str_to_title(class),
         n_am_tot = n()) %>%        ### length of whole AM maps list
  group_by(class) %>%
  summarize(n_am_class_all  = n(),
            p_am_class_all  = round((n_am_class_all/first(n_am_tot) * 100), 2))

taxa_am_solo <- spp_am %>%
  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
  filter(!sciname %in% iucn_maps_class$sciname) %>%  ### filter list to AM only
  mutate(class = str_to_title(class),
         n_am_solo = n()) %>%       ### length of AM only list
  group_by(class) %>%
  summarize(n_am_class_solo = n(),
            p_am_class_solo = round((n_am_class_solo/first(n_am_solo) * 100), 2))

taxa_both <- spp_am %>%
  filter(sciname %in% spp_iucn_maps$sciname) %>%
  filter(!class %in% c('Not assigned', 'null')) %>%
  mutate(class = str_to_title(class),
         n_both_tot = n()) %>%
  group_by(class) %>%
  summarize(n_class_both = n(),
            p_class_both = round(n_class_both/first(n_both_tot) * 100, 2))

taxa_df <- taxa_am_all %>%
  full_join(taxa_am_solo,   by = 'class') %>%
  full_join(taxa_iucn_all,  by = 'class') %>%
  full_join(taxa_iucn_solo, by = 'class') %>%
  full_join(taxa_both,      by = 'class') %>%
  filter(!(is.na(n_iucn_class_all) & p_am_class_all < 0.5))

knitr::kable(taxa_df %>%
               select(class, am_perc = p_am_class_all, iucn_perc = p_iucn_class_all, both_perc = p_class_both))
```

## Spatial global distribution between data sets

# Analysis of spatial alignment by species

## Defining spatial alignment between the two data sets

@@

## Comparing spatial alignment between the two data sets

@@

___Figure XXX: Distribution alignment vs area alignment for `r nrow(spp_map_pairs)` species mapped in both Aquamaps and IUCN species distribution maps.___

___ditch the species identifiers here?  or highlight ones that are used in discussion of possible mechanistic explanations for differences?___

@@

@@

___move to discussion?___ We were not surprised to find that map pairs for marine mammals, tunas, and billfishes were well-matched (quadrants 1 and 2); many of these species are global-ranging pelagic organisms, with some ranges encompassing nearly the entire ocean.  Mangroves are also not a surprise, being immobile and easily mapped by remote sensing methods.

@@

___Other hypotheses/case studies to highlight possible mechanistic differences?___

# AquaMaps: Effect of changing "presence" threshold on area representation

![](figures/AMThreshold.png) ___add a mean/median vline on this?___ do we need this figure? it looks cool though... why broken down by risk category? Ditch NE category since most studies probably care about the species health as well as distribution?

___Identify several studies that used different thresholds - OHI, MPA gap analysis, AquaMaps mammal distribution___ to justify this micro-analysis

![](figures/scatter_AMThreshold_mockup.png)
Effect of selecting different thresholds on total "present" area - linearish, no big steps or drops to indicate sudden change in tradeoff.  Do regression - shallow slope indicates that sensitivity is less than 1:1.

@@
